### 2019-04-01 2-Chapters/2-Chapter/chapter2.tex
Still fighting against my template...
Okay for chapters and parts
Small modifications on all chapters

% First chapter begins here
\chapter{Single Player and Stationary Multi-Armed Bandit Models}
\label{chapter:2}
\minitoc
\newpage
% Write miniTOC just after the title
\graphicspath{{2-Chapters/2-Chapter/Images/}}

@@

% ----------------------------------------------------------------------------
\section{Different approaches on algorithm selection, or ``How to choose your preferred bandit algorithm?''}
\label{sec:2:chooseYourPreferredBanditAlgorithm}

@@

A last solution is online algorithm selection, inspired from expert aggregation.
Include here the discussion about expert aggregation and my \textbf{Aggregator} algorithm, see https://hal.inria.fr/hal-01705292

>>>>

% First chapter begins here
\chapter{Single Player and Stationary Multi-Armed Bandit Models}
\label{chapter:2}
\minitoc

Abstract for Chapter 2.

\newpage
% Write miniTOC just after the title
\graphicspath{{2-Chapters/2-Chapter/Images/}}

@@

% ----------------------------------------------------------------------------
\section{Different approaches on algorithm selection}
\label{sec:2:chooseYourPreferredBanditAlgorithm}

@@

% A last solution is online algorithm selection, inspired from expert aggregation.
% Include here the discussion about expert aggregation and my \textbf{Aggregator} algorithm, see https://hal.inria.fr/hal-01705292

### 2019-04-01 2-Chapters/3-Chapter/chapter3.tex
Small modifications on all chapters

SMPyBandits is a package I have developed since the beginning of my PhD.
It is designed to allow easy and efficient numerical simulations on single-player and multi-players Multi-Armed Bandits (MAB) algorithms, and SMPyBandits is written in Python.
This library is by far the most complete open-source implementation of state-of-the-art algorithms tackling various kinds of MAB sequential learning problems.
It aims at being extensive, simple to use and maintain, with a clean and well documented codebase.
It allows fast prototyping of simulations and experiments, with an easy configuration system and command-line options to customize experiments.
Experiment results are saved in an optimized binary format (HDF5) as well as high quality plots of many useful visualizations.
%
More than two years of active development have shown how easy it can be to add new algorithms, new arm distributions, and new bandit models (\eg, Markovian or non-stationary).

This chapter details the organization of the library, what it implements in terms of arm distributions, models, algorithms and visualizations.
Then we use it to perform a numerical comparisons of the main state-of-the-art single-player MAB algorithms, as well as a study to compare time and memory costs of the main algorithms.
Finally, we introduce rested or restless Markov models and use the library to simulate and illustrate the performance of the classical \UCB{} algorithm.

\newpage
% Write miniTOC just after the title
\graphicspath{{2-Chapters/3-Chapter/Images/}}
\graphicspath{{2-Chapters/3-Chapter/SMPyBandits_paper.git/plots/}}

@@

SMPyBandits does not aim at being blazing fast or perfectly memory efficient, and comes with a pure Python implementation \cite{python}, with no dependency except standard open-source Python packages.
Even if some critical parts are also available as a \texttt{C} Python extension, and even by using Numba \cite{numba} whenever it is possible, if simulation speed really matters, one should rather refer to less exhaustive but faster implementations, like for example \cite{TorLibbandit} in \texttt{C++} or \cite{VishMABjl} in Julia.

@@

In \cite{Besson2018WCNC} (see Chapter~\ref{chapter:2}), we used SMPyBandits to illustrate and compare different aggregation algorithms\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/Aggregation.html}{SMPyBandits.GitHub.io/Aggregation.html}} on the documentation.}. We designed a variant of the Exp3 algorithm for online aggregation of experts \cite{Bubeck12}, called \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Aggregator.html}{Aggregator}}. Aggregating experts is a well-studied idea in sequential learning and in machine learning in general. We showed that it can be used in practice to select on the run the best bandit algorithm for a certain problem from a fixed pool of experts. This idea and algorithm can have interesting impact for Opportunistic Spectrum Access applications \cite{Jouini09} that use multi-armed bandits algorithms for sequential learning and network efficiency optimization.

- For \cite{Besson2018ALT} (see Chapter~\ref{chapter:5}), we used SMPyBandits for all the simulations for multi-player bandit algorithms\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/MultiPlayers.html}{SMPyBandits.GitHub.io/MultiPlayers.html}} on the documentation.}. We designed the two \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.RandTopM.html}{RandTopM}} and \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.MCTopM.html}{MCTopM}} algorithms and proved than they enjoy logarithmic regret in the usual setting, and outperform significantly the previous state-of-the-art solutions (\ie, \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.rhoRand.html}{rhoRand}}, \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MEGA.html}{MEGA}} and \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MusicalChair.html}{MusicalChair}}).

In \cite{Besson2018DoublingTricks}, we used SMPyBandits to illustrate and compare different "doubling trick" schemes\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/DoublingTrick.html}{SMPyBandits.GitHub.io/DoublingTrick.html}} on the documentation.}. In sequential learning, an algorithm is anytime if it does not need to know the horizon $T$ of the experiments. A well-known trick for transforming any non-anytime algorithm to an anytime variant is the ``Doubling Trick'': start with an horizon $T_0\in\mathbb{N}^*$, and when $t > T_i$, use $T_{i+1} = 2 T_i$. We studied two generic sequences of growing horizons (geometric and exponential), and we proved two theorems that generalized previous results. A geometric sequence suffices to conserve minimax regret bounds (in $R_T = \mathcal{O}(\sqrt{T})$), with a constant multiplicative loss $\ell \leq 4$, but cannot be used to conserve a logarithmic regret bound (in $R_T = \mathcal{O}(\log(T))$). And an exponential sequence can be used to conserve logarithmic bounds, with a constant multiplicative loss also $\ell \leq 4$ in the usual setting. It is still an open question to know if a well-tuned exponential sequence can conserve minimax bounds, or "weak" minimax bounds (in $R_T = \mathcal{O}(\sqrt{T \log(T)})$).

In \cite{Besson2019GLRT} (see Chapter~\ref{chapter:6}), we used SMPyBandits for piece-wise stationary MAB models (single-player). We illustrate and compare different algorithms designed for this family of non-stationary problems\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/NonStationary.html}{SMPyBandits.GitHub.io/NonStationary.html}} on the documentation.}. XXX

\TODOL{Write a better explanation of this last paper.}

>>>>

SMPyBandits is a package I have developed since the beginning of my PhD.
It is designed to allow easy and efficient numerical simulations on single-player and multi-players Multi-Armed Bandits (MAB) algorithms, and SMPyBandits is written in Python.
This library is by far the most complete open-source implementation of state-of-the-art algorithms tackling various kinds of MAB sequential learning problems.
It aims at being extensive, simple to use and maintain, with a clean and well documented codebase.
It allows fast prototyping of simulations and experiments, with an easy configuration system and command-line options to customize experiments.
Experiment results are saved in an optimized binary format (HDF5) as well as high quality plots of many useful visualizations.
%
More than two years of active development have shown how easy it can be to add new algorithms, new arm distributions, and new bandit models (\eg, Markov or non-stationary).

This chapter details the organization of the library, what it implements in terms of arm distributions, models, algorithms and visualizations.
Then we use it to perform a numerical comparisons of the main state-of-the-art single-player MAB algorithms, as well as a study to compare time and memory costs of the main algorithms.
\TODOL{Do we really have time/space to talk about Markov models?}
Finally, we introduce rested or restless Markov models and use the library to simulate and illustrate the performance of the classical \UCB{} algorithm.

% \newpage
% Write miniTOC just after the title
\graphicspath{{2-Chapters/3-Chapter/Images/}}
\graphicspath{{2-Chapters/3-Chapter/SMPyBandits_paper.git/plots/}}

@@

SMPyBandits does not aim at being blazing fast or optimal in terms of memory usage, and it comes with a pure Python implementation \cite{python}, with no dependency except standard open-source Python packages.
Even if some critical parts are also available as a \texttt{C} Python extension, and even by using Numba \cite{numba} whenever it is possible, if simulation speed really matters, one should rather refer to less exhaustive but faster implementations, like for example \cite{TorLibbandit} in \texttt{C++} or \cite{VishMABjl} in Julia.

@@

In \cite{Besson2018WCNC} and in Chapter~\ref{chapter:2}, we used SMPyBandits to illustrate and compare different aggregation algorithms\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/Aggregation.html}{SMPyBandits.GitHub.io/Aggregation.html}} on the documentation, which gives detailed instructions to reproduce the results presented in this research paper, details about the models and the notations, and bibliographic references.}. We designed a variant of the Exp4 algorithm for online aggregation of experts \cite{Bubeck12}, called \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Aggregator.html}{Aggregator}}.
% Aggregating experts is a well-studied idea in sequential learning and in machine learning in general. We showed that it can be used in practice to select on the run the best bandit algorithm for a certain problem from a fixed pool of experts. This idea and algorithm can have interesting impact for Opportunistic Spectrum Access applications \cite{Jouini09} that use multi-armed bandits algorithms for sequential learning and network efficiency optimization.

In \cite{Besson2018ALT} and in Chapter~\ref{chapter:5}, we used SMPyBandits for all the simulations for multi-player bandit algorithms\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/MultiPlayers.html}{SMPyBandits.GitHub.io/MultiPlayers.html}} on the documentation.}. We designed the two \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.RandTopM.html}{RandTopM}} and \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.MCTopM.html}{MCTopM}} algorithms and proved than they enjoy logarithmic regret in the usual setting, and outperform significantly the previous state-of-the-art solutions (\ie, \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.rhoRand.html}{rhoRand}}, \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MEGA.html}{MEGA}} and \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MusicalChair.html}{MusicalChair}}).

In \cite{Besson2018DoublingTricks}, we used SMPyBandits to illustrate and compare different "doubling trick" schemes\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/DoublingTrick.html}{SMPyBandits.GitHub.io/DoublingTrick.html}} on the documentation.}. and lots of bibliographic referencesIn sequential learning, an algorithm is anytime if it does not need to know the horizon $T$ of the experiments. A well-known trick for transforming any non-anytime algorithm to an anytime variant is the ``Doubling Trick'': start with an horizon $T_0\in\mathbb{N}^*$, and when $t > T_i$, use $T_{i+1} = 2 T_i$. We studied two generic sequences of growing horizons (geometric and exponential), and we proved two theorems that generalized previous results. A geometric sequence suffices to conserve minimax regret bounds (in $R_T = \mathcal{O}(\sqrt{T})$), with a constant multiplicative loss $\ell \leq 4$, but cannot be used to conserve a logarithmic regret bound (in $R_T = \mathcal{O}(\log(T))$). And an exponential sequence can be used to conserve logarithmic bounds, with a constant multiplicative loss also $\ell \leq 4$ in the usual setting. It is still an open question to know if a well-tuned exponential sequence can conserve minimax bounds, or "weak" minimax bounds (in $R_T = \mathcal{O}(\sqrt{T \log(T)})$).

In \cite{Besson2019GLRT} and \cite{Besson2019Gretsi}, and in Chapter~\ref{chapter:6}, we used SMPyBandits for piece-wise stationary MAB models (only for the single-player case). We illustrate and compare different algorithms designed for this family of non-stationary problems\footnote{~See the page \texttt{\href{https://SMPyBandits.GitHub.io/NonStationary.html}{SMPyBandits.GitHub.io/NonStationary.html}} on the documentation.}.
We designed the \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.GLR_UCB.html}{GLRklUCB}} policy, and implemented most of the policies from state-of-the-art research on passively or active adaptive policies, both adversarial- or stochastic-based, designed to tackle the piece-wise stationary MAB problems.
We also implemented a large benchmark of problems.

### 2019-04-01 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Small modifications on all chapters

\begin{figure}[!b]
    \centering
    \subfloat[10\% of dynamic devices]{\includegraphics[scale=0.5]{10intelligent.eps}
    \label{fig:41:10intelligent}}
    \hfill
    \subfloat[30\% of dynamic devices]{\includegraphics[scale=0.5]{30intelligent.eps}
    \label{fig:41:30intelligent}}
    \vskip\baselineskip
    \vspace*{-20pt}
    \subfloat[50\% of dynamic devices]{\includegraphics[scale=0.5]{50intelligent.eps}
    \label{fig:41:50intelligent}}
    \hfill
    \subfloat[100\% of dynamic devices]{\includegraphics[scale=0.5]{100intelligent.eps}
    \label{fig:41:100intelligent}}
    \caption{Performance of two MAB algorithms (\UCB{} and Thompson Sampling), compared to extreme references without learning or oracle knowledge, when the proportion of dynamic end-devices in the network increases, from $10\%$ to $100\%$.}
    \label{fig:41:from10to100}
    \vspace*{-10pt}
\end{figure}

>>>>

\begin{figure}[!b]
    \centering
    \subfloat[10\% of dynamic devices]{\includegraphics[scale=0.5]{10intelligent.eps}
    \label{fig:41:10intelligent}}
    \hfill
    \subfloat[30\% of dynamic devices]{\includegraphics[scale=0.5]{30intelligent.eps}
    \label{fig:41:30intelligent}}
    \vskip\baselineskip
    \vspace*{-20pt}
    \subfloat[50\% of dynamic devices]{\includegraphics[scale=0.5]{50intelligent.eps}
    \label{fig:41:50intelligent}}
    \hfill
    \subfloat[100\% of dynamic devices]{\includegraphics[scale=0.5]{100intelligent.eps}
    \label{fig:41:100intelligent}}
    % FIXME use a shorter caption for the list of figures in Appendix of the thesis: it takes lots of lines to have these (a)...(d) shown in list of figures
    \caption[Performance of two MAB algorithms (\UCB{} and Thompson Sampling), compared to extreme references without learning or oracle knowledge, when the proportion of dynamic end-devices in the network increases, from $10\%$ to $100\%$.]{Performance of two MAB algorithms (\UCB{} and Thompson Sampling), compared to extreme references without learning or oracle knowledge, when the proportion of dynamic end-devices in the network increases, from $10\%$ to $100\%$.}
    \label{fig:41:from10to100}
    \vspace*{-10pt}
\end{figure}

### 2019-04-01 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Small modifications on all chapters

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.85\textwidth]{UI.eps}
    \caption{User interface of our demonstration.}
    \label{fig:42:UI}
\end{figure}

@@

\paragraph{Video}
%
A $6$-minute \textbf{video} showing our demonstration is available online at \texttt{\url{youtu.be/HospLNQhcMk}}.
It shows examples of $3$ dynamic devices learning simultaneously, confirming the results of Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} for overlapping channels.
It also shows the connections between the USRP boards, the Octoclock, the master laptop etc, completing the presentation of the SCEE testbed already shown in Figure~\ref{fig:42:photosSCEETestBed}.

>>>>

\begin{figure}[!h]
    \centering
    \includegraphics[width=1.00\textwidth]{UI.eps}
    \caption{User interface of our demonstration.}
    \label{fig:42:UI}
\end{figure}

@@

\paragraph{Video}
%
As shown below in Figure~\ref{fig:42:screenshotDemoYouTube},
we made a $6$-minute \textbf{video} to show our demonstration online, and it is available on the YouTube hosting platform, at \texttt{\href{https://youtu.be/HospLNQhcMk}{youtu.be/HospLNQhcMk}}.
The video shows examples of $3$ dynamic devices learning simultaneously, confirming the results of Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} for overlapping channels.
It also shows the connections between the USRP boards, the Octoclock, the master laptop etc, completing the presentation of the SCEE testbed already shown in Figure~\ref{fig:42:photosSCEETestBed}.

\begin{figure}[!h]
	\centering
    % \includegraphics[width=0.90\textwidth]{Images/screenshotDemoYouTube.png}
    \includegraphics[width=0.90\textwidth]{2-Chapters/4-Chapter/Images/screenshotDemoYouTube.png}
    \caption{Screenshot of the \textbf{video} of our demonstration, \texttt{\href{https://youtu.be/HospLNQhcMk}{youtu.be/HospLNQhcMk}}.}
    \label{fig:42:screenshotDemoYouTube}
\end{figure}

### 2019-04-01 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main_appendix.tex
Done for now for Chapter4
Small modifications on all chapters

\begin{figure}[!b]
	\centering
    \includegraphics[width=1.00\textwidth]{2-Chapters/4-Chapter/Images/USRP_TX_PU__v1__simple_grc.png}
    \caption{The \textbf{random traffic generator}. \texttt{Generator} is the only hand-written block, which is configured with a list of active channels, a list of occupancy rate, and constants about the PHY layer (preamble length and data length). It uses one USRP only for emitting data (``Sink'' mode).}
    \label{fig:4app:USRP_TX_PU__v1__simple_grc}
\end{figure}

\begin{figure}[!b]
	\centering
    \includegraphics[width=1.00\textwidth]{2-Chapters/4-Chapter/Images/USRP_RX_BTS__v1__simple_grc.png}
    \caption{The \textbf{IoT gateway}. The hand-written blocks are the 1) \texttt{Demodulator} block, which is configured with a list of detection threshold (on the received power), 2) the \texttt{Check\_ack} block, which is configured with a maximum block error and an accepted error rates (to decide when a message is close enough to an acknowledgement), and finally 3) the \texttt{send\_ack} block, which is configured with the list of active channels, and knowledge about the PHY layer (delay before sending the Ack and its length). All blocks need to know constants about the PHY layer (preamble length and data length). The IoT gateway uses one USRP for both emitting and receiving data (``Sink'' and ``Source'' modes).}
    \label{fig:4app:USRP_RX_BTS__v1__simple_grc}
\end{figure}

\begin{figure}[!b]
	\centering
    \includegraphics[width=1.00\textwidth]{2-Chapters/4-Chapter/Images/USRP_TX_SU__v1__simple_grc.png}
    \caption{The \textbf{IoT dynamic object}. The hand-written blocks are the 1) the \texttt{Renormalized\_ack} block, which is configured with the list of active channels and extra knowledge about the PHY layer (preamble length, message length, and a threshold to tune detection of the incoming Ack), 2) the \texttt{Demodulator} and 3) the \texttt{Check\_ack} blocks, both shared with the IoT gateway, and finally 4) the \texttt{generator\_SU} block, which embeds the \UCB{} or Thompson Sampling algorithm, and which is configured with the list of active channels. Most blocks need to know constants about the PHY layer (preamble length and data length). Any of the IoT dynamic objects use one USRP for both emitting and receiving data (``Sink'' and ``Source'' modes).}
    \label{fig:4app:USRP_TX_SU__v1__simple_grc}
\end{figure}

>>>>

\begin{figure}[!b]
	\centering
    \includegraphics[width=1.00\textwidth]{2-Chapters/4-Chapter/Images/USRP_TX_PU__v1__simple_grc.png}
    \caption[The \textbf{random traffic generator} flow-graph.]{The \textbf{random traffic generator}. \texttt{Generator} is the only hand-written block, which is configured with a list of active channels, a list of occupancy rate, and constants about the PHY layer (preamble length and data length). It uses one USRP only for emitting data (``Sink'' mode).}
    \label{fig:4app:USRP_TX_PU__v1__simple_grc}
\end{figure}

\begin{figure}[!b]
	\centering
    \includegraphics[width=1.00\textwidth]{2-Chapters/4-Chapter/Images/USRP_RX_BTS__v1__simple_grc.png}
    \caption[The \textbf{IoT gateway} flow-graph.]{The \textbf{IoT gateway}. The hand-written blocks are the 1) \texttt{Demodulator} block, which is configured with a list of detection threshold (on the received power), 2) the \texttt{Check\_ack} block, which is configured with a maximum block error and an accepted error rates (to decide when a message is close enough to an acknowledgement), and finally 3) the \texttt{send\_ack} block, which is configured with the list of active channels, and knowledge about the PHY layer (delay before sending the Ack and its length). All blocks need to know constants about the PHY layer (preamble length and data length). The IoT gateway uses one USRP for both emitting and receiving data (``Sink'' and ``Source'' modes).}
    \label{fig:4app:USRP_RX_BTS__v1__simple_grc}
\end{figure}

\begin{figure}[!b]
	\centering
    \includegraphics[width=1.00\textwidth]{2-Chapters/4-Chapter/Images/USRP_TX_SU__v1__simple_grc.png}
    \caption[The \textbf{IoT dynamic device} flow-graph.]{The \textbf{IoT dynamic device}. The hand-written blocks are the 1) the \texttt{Renormalized\_ack} block, which is configured with the list of active channels and extra knowledge about the PHY layer (preamble length, message length, and a threshold to tune detection of the incoming Ack), 2) the \texttt{Demodulator} and 3) the \texttt{Check\_ack} blocks, both shared with the IoT gateway, and finally 4) the \texttt{generator\_SU} block, which embeds the \UCB{} or Thompson Sampling algorithm, and which is configured with the list of active channels. Most blocks need to know constants about the PHY layer (preamble length and data length). Any of the IoT dynamic devices use one USRP for both emitting and receiving data (``Sink'' and ``Source'' modes).}
    \label{fig:4app:USRP_TX_SU__v1__simple_grc}
\end{figure}
